STR: Seasonal-Trend decomposition using Regression

### 2021-03-25 Response_to_reviewers.Rmd

---
title: Response to reviewers
author: "STR: Seasonal-Trend decomposition using Regression"
bibliography: strrefs.bib
output: MonashEBSTemplates::memo
header-includes:
  - \usepackage{color}
  - \renewenvironment{quote}{\list{}{\rightmargin\leftmargin}\item\relax\color[RGB]{0,150,0}}{\endlist}

---

> We thank the reviewers for their careful reading of our paper; their comments have led to several improvements and corrections. Reviewer comments are in black, our responses are in green.

# Senior Editor

I join the AE in thanking the authors for their submission. The manuscript presents an innovative approach to an important problem in an approachable way that I expect to be broadly interesting -- and impactful -- to both researchers and practitioners in the data science community. This definitely fits the template for inclusion in an early volume of IJDS.

We conclude that the apparent promise of the proposed algorithm has not been computationally demonstrated in a thorough enough manner to warrant publication. The specific suggestions of the reviewer have been clearly summarized by the AE, and at least the first three categories outlined by the AE should be required for publication. This will necessitate some significant computational experiments and careful analysis of results. Still, since the conceptual case for the method has been sufficiently made, and the description (and code) are sufficient for reproducibility, the resulting recommendation is "minor revision."

We look forward to reviewing a revised version of the manuscript.

# Associate Editor

I would like to thank the authors for having the opportunity to review their innovative work on decomposing complex time series data. The paper has been reviewed by one referee who has made valuable suggestions. I will, in what follows, summarize some of these suggestions:

- “Does it work?” When we read a new research contribution, we often ask ourselves “does this method work?” That is, we ask ourselves that, while the authors propose a new method to recover underlying or hidden time series components, does the method really uncover all of the components hidden in the data? In this paper, the authors provide some empirical examples, but they do not really answer the question of whether all the time series components, of e.g. the electricity demand in Victoria, have been truly recovered. What if there are additional components that are hiding inside the data (which the method could not detect)? In order to address this question, the referee suggests the use of simulations. Simulations are a great tool to investigate what-if questions (e.g. what if we added an hourly or minutely component to the data? Would the method then still work? What if we added more noise to the data? Or unobservable factors?) I would strongly recommend the authors consider the referee’s suggestion and see if a simulation experiment could be added to the paper.

> As suggested, we have added some simulation experiments to demonstrate our proposed method. These appear in Section ?? and in the appendix.

- “Does it work better?” Another question we often ask ourselves is the “so what” question. Or more specifically, we ask if the proposed method, while elegant and convincing, really adds any practical value. In other words, does it work better than any of the existing methods? In the empirical examples, while the authors show the performance of STR, they do not show us how poorly (or how competitively) other benchmark methods perform. The authors argue that none of the existing methods satisfy all of the requirements of STR; however, I agree with the referee that it would add tremendous value if some benchmarking could be added to the empirical applications. That would help the reader better understand how much better STR performs (and when it performs better).

> We have now included some comparison methods in the simulations and in the appendix.

- Special cases: It appears that a very special case of the model would be a linear regression model with a linear trend and seasonal dummy variables (plus covariates). I think it would have tremendous value for the reader if the authors could discuss special cases of their model and how they relate to the standard statistics and economics tool sets.

> Yes, that is correct. We have added a paragraph after the introduction of our model describing some special cases.

- Applications within data science: One way to look at this method of time series decomposition is in terms of feature engineering. The authors’ model provides a way for the data science researcher to identify new features in the data (such as weekly or daily patterns). Could these features then be used inside other data science methods (e.g. could the uncovered weekly or daily pattern be used as an input for a classification or forecasting model)? I think it would provide tremendous value for the data science community if the authors could discuss the application of their model/method within the wider context of data science tasks.

# Reviewer

General Assessment: The paper is very well written, well organized, and easy to follow. I believe the proposed method to be of interest to the target audience of the IJDS. However, I think the paper presents a few loose ends. Let me elaborate with a bullet-point style to facilitate the authors' evaluation.

1) Identification: I don't think the authors are convincing the reader whether the proposed method can recover unbiased parameters. I suggest the authors include extensive simulation results, either in the main text or in a supplementary appendix. Using simulations would also help assess what kind of "richness" in the data is needed to correctly estimate the model; this is an essential requirement for a data science paper.

2) Benchmarking: the authors should include potential alternative (traditional STR method) or multiscale/wavelet alike (See Zhang et al. 2017 for a quick reference and potential suggestion) method and compare the predictive performance across them.

3) I suggest the authors consider an additional dataset that could make the paper more appealing to a vaster audience. As a suggestion, the authors could consider time series extracted from Google Trends about Covid (https://trends.google.com/trends/story/US_cu_4Rjdh3ABAABMHM_en)

4) I enjoyed reading the extended yet brief discussion in Section 4. I think that 4.2 could be implemented in the main paper to make a more substantial methodological contribution. It would be indeed intriguing to see how the bias/variance trade-off obtained using different regularizers (or a combination thereof) can impact the results' quality.

Overall, as remarked above, it is quite an enjoyable paper with room to improve. I trust the authors find my comment useful under any outcome.

### 2021-03-28 str_ijds.Rmd
Simulations added

---
title: "STR: Seasonal-Trend decomposition using Regression"
author:
 - familyname: Dokumentov
   othernames: Alexander
   address: Growing Data\newline Australia
 - familyname: Hyndman
   othernames: Rob J
   address: Monash University\newline Australia
   email: Rob.Hyndman@monash.edu
   correspondingauthor: true
abstract: "We propose a new method for decomposing seasonal data: STR (a Seasonal-Trend decomposition using Regression). Unlike other decomposition methods, STR allows for multiple seasonal and cyclic components, covariates, seasonal patterns that may have non-integer periods, and seasonality with complex topology. It can be used for time series with any regular time index including hourly, daily, weekly, monthly or quarterly data. It is competitive with existing methods when they exist, but tackles many more decomposition problem than other methods allow.\\par STR is based on a regularized optimization, and so is somewhat related to ridge regression. Because it is based on a statistical model, we can easily compute confidence intervals for components, something that is not possible with most existing decomposition methods (such as STL, X-12-ARIMA, SEATS-TRAMO, etc.).\\par
  Our model is implemented in the R package *stR*, so can be applied by anyone  to their own data."
keywords: "time series decomposition, seasonal data, Tikhonov regularisation, ridge regression, LASSO, STL, TBATS, X-12-ARIMA, BSM"
blind: false
cover: false
toc: false
bibliography: strrefs.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    toc: no
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
header-includes:
  - \graphicspath{{figs/}}
  - \def\addlinespace{}
  - \usepackage{amsmath}
  - \setlength{\arraycolsep}{3pt}
  - \setcounter{MaxMatrixCols}{20}
  - \newcommand{\argmin}{\operatorname{arg\,min}}
  - \newcommand{\sn}{\kappa}
  - \renewcommand{\vec}{\operatorname{vec}}
  - \newcommand{\diag}{\operatorname{diag}}
  - \renewcommand{\mod}{\operatorname{mod}}
  - \usepackage[section]{placeins}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  cache=TRUE
)
library(tidyverse)
library(stR)
library(fabletools)
library(distributional)
library(lubridate)

# Create predictor dist object
predictor_dist <- function(x, level) {
    dist_percentile(
      with(x, mapply(c, data, split(lower, row(lower)), split(upper, row(upper)), SIMPLIFY = FALSE)),
      list(c(50, 50 + rep(c(-1, 1), each = length(level))*rep(level, 2)/2))
    )
  }
```

@@

@@

```{r supermarket, fig.height=7, fig.cap="Supermarket and grocery store log revenue in New South Wales, decomposed with STR. The original data (logged) are shown in the top panel, while the components are shown in the other panels, with 95\\% confidence intervals for the trend and season components in blue"}
supermarket_nsw <- tsibbledata::aus_retail %>%
  filter(
    State == "New South Wales",
    Industry == "Supermarket and grocery stores",
    year(Month) >= 2000,
    year(Month) <= 2009
  ) %>%
  pull(Turnover) %>%
  ts(start=2000, frequency=12)

decomp <- log(supermarket_nsw) %>%
  AutoSTR(confidence=0.95, gapCV=12, reltol=0.00001)

as_tsibble(decomp$input$data) %>%
  rename(log_Data = value) %>%
  mutate(
    Remainder = decomp$output$random$data,
    Trend = predictor_dist(decomp$output$predictors[[1]], 95),
    Season = predictor_dist(decomp$output$predictors[[2]], 95)
  ) %>%
  as_dable(response = "log_Data", aliases = rlang::exprs(log_Data = Trend + Season + Remainder), method="STR") %>%
  autoplot(size=0.5, fill='blue', level=95) +
  theme(legend.position = "none") +
  xlab("Month") +
  theme_bw() +
  theme(legend.position = "none")
```

@@

This data set has two seasonal patterns. The first pattern is a weekly seasonal pattern ($m_1=48\times7=336$) that represents the specific demand features that are attributable to a particular day of the week. The second pattern is a daily seasonal pattern ($m_2=48$) with the topology in Figure \ref{fig:topology2_pdf}, which allows the model to distinguish between working days and holidays/weekends, and to make transitions between them. The pattern reflects the tendency to have a higher electricity demand during standard working hours and a lower demand at night. It also reflects the tendency to have different demand patterns on working days and holidays/weekends. A longer series would also have an annual seasonal pattern, but with only 115 days, this cannot be distinguished from the trend component.

```{r electricity, cache=FALSE}
# Read and join all data sets
elec <- read_csv("data/VIC.csv") %>%
    mutate(Date = as.Date(Date, origin = "1899-12-30")) %>%
    left_join(
      read_csv("data/Melbourne.csv") %>%
        mutate(Date = as.Date(Date, origin = "1899-12-30")),
      by = c("Date", "Period")
    ) %>%
    left_join(
      read_csv("data/VIC public holidays.csv", quote = "\"", col_names = "Date") %>%
        mutate(
          Date = lubridate::dmy(Date),
          Holidays = TRUE
        ),
      by = c("Date")
    ) %>%
    replace_na(list(Holidays = FALSE)) %>%
    rename(Demand = OperationalLessIndustrial) %>%
    select(-Industrial) %>%
    mutate(Time = seq_along(Date)) %>%
    filter(Date >= "2000-01-11", Date <= "2000-05-04") %>%
    # Add in day, week and working day info
    mutate(
      Weekday = factor(weekdays(Date, abbreviate = TRUE),
        levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
      DailySeasonality = Period - 1,
      WeeklySeasonality = (as.integer(Weekday) - 1) * 48 + Period - 1,
      WorkingDay = !Holidays & Weekday %in% c("Mon", "Tue", "Wed", "Thu", "Fri"),
      WDSeasonality = (DailySeasonality + (!WorkingDay) * 100),
      DateTime = as.POSIXct(paste(Date, "00:00:00")) + (Period-1)*30*60
    )

# Have we already done this?
if(file.exists("data/electricity_str.rds")) {
  elec_str <- read_rds("data/electricity_str.rds")
} else {
  # Set up predictors for STR
  Predictors <- list()
  Predictors$Trend <- list(
    name = "Trend",
    data = rep(1, NROW(elec)),
    times = elec$Time,
    seasons = rep(1, NROW(elec)),
    timeKnots = seq(from = first(elec$Time), to = last(elec$Time), length.out = 116),
    seasonalStructure = list(segments = list(c(0, 1)), sKnots = list(c(1, 0))),
    lambdas = c(1500, 0, 0)
  )
  Predictors$WSeason <- list(
    name = "Weekly seas",
    data = rep(1, NROW(elec)),
    times = elec$Time,
    seasons = elec$WeeklySeasonality,
    timeKnots = seq(from = first(elec$Time), to = last(elec$Time), length.out = 12),
    seasonalStructure = list(segments = list(c(0, 336)), sKnots = c(as.list(seq(4, 332, 4)), list(c(336, 0)))),
    lambdas = c(0.8, 0.6, 100)
  )
  Predictors$WDSeason <- list(
    name = "Daily seas",
    data = rep(1, NROW(elec)),
    times = elec$Time,
    seasons = elec$WDSeasonality,
    timeKnots = seq(from = first(elec$Time), to = last(elec$Time), length.out = 24),
    seasonalStructure = list(
      segments = list(c(0, 48), c(100, 148)),
      sKnots = c(as.list(c(1:47, 101:147)), list(c(0, 48, 100, 148)))
    ),
    lambdas = c(0.003, 0, 240)
  )
  Predictors$TrendTempM <- list(
    name = "Trend temp",
    data = elec$Temp,
    times = elec$Time,
    seasons = rep(1, NROW(elec)),
    timeKnots = Predictors$Trend$timeKnots,
    seasonalStructure = Predictors$Trend$seasonalStructure,
    lambdas = c(1e7, 0, 0)
  )
  Predictors$TrendTempM2 <- list(
    name = "Trend temp^2",
    data = elec$Temp^2,
    times = elec$Time,
    seasons = rep(1, NROW(elec)),
    timeKnots = Predictors$Trend$timeKnots,
    seasonalStructure = Predictors$Trend$seasonalStructure,
    lambdas = c(1e7, 0, 0)
  )

@@

```{r vic, fig.height=10, fig.cap="Electricity demand in Victoria, decomposed using STR. The original data are plotted in black in the first panel, with the decomposed trend plotted in red. The subsequent panels contain the weekly seasonal pattern; the daily seasonal pattern with complex topology; the effect of temperature in Melbourne, which has a daily seasonal pattern with complex topology; the effect of squared temperatures in Melbourne, which is allowed to vary over time; and the residuals. Grey vertical lines represent the residuals with the largest absolute values.", dependson='outliers'}
elec_dbl <- elec %>%
  as_tsibble(index=DateTime) %>%
  mutate(
    Remainder = elec_str$output$random$data,
    Trend = predictor_dist(elec_str$output$predictors[[1]], 95),
    Weekly = predictor_dist(elec_str$output$predictors[[2]], 95),
    Daily = predictor_dist(elec_str$output$predictors[[3]], 95),
    Temp = predictor_dist(elec_str$output$predictors[[4]], 95),
    Tempsq = predictor_dist(elec_str$output$predictors[[5]], 95),
  ) %>%
  select(DateTime, Demand, Trend, Weekly, Daily, Temp, Tempsq, Remainder) %>%
  as_dable(response = "Demand",
           aliases = rlang::exprs(Demand = Trend + Weekly + Daily + Temp + Tempsq + Remainder),
           method = "STR")

p <- elec_dbl %>%
  autoplot(size = 0.5, fill='blue', level=95) +
  xlab("Month") +
  geom_vline(data=outliers, aes(xintercept = DateTime), col='gray') +
  ggplot2::theme_bw() + theme(legend.position = "none")
# Put vline behind data
p$layers <- p$layers[c(1,4,2,3)]
p
```

@@

```{r outlierstable, results='asis'}
outliers %>%
  mutate(
    Hour1 = hour(DateTime),
    Hour2 = hour(DateTime+1800),
    Minute1 = minute(DateTime) %>% as.character(),
    Minute2 = minute(DateTime+1800) %>% as.character(),
    Minute1 = if_else(Minute1=="0","00",Minute1),
    Minute2 = if_else(Minute2=="0","00",Minute2),
    Time = paste0(Hour1,":",Minute1," -- ",Hour2,":",Minute2),
    Weekday = case_when(
      Weekday == "Thu" ~ "Thursday",
      Weekday == "Fri" ~ "Friday"
    ),
    Date = paste(day(Date),month(Date, label=TRUE, abbr=FALSE),year(Date)),
  ) %>%
  select(Date,Weekday,Time,res) %>%
  knitr::kable(col.names = c("Date","Day of week","Time period","Residual"),
               digits=1, align="rlcr", booktabs=TRUE,
               caption = "The ten residuals that are largest in absolute value after an STR decomposition."
  )
```

```{r MelTemp, dependson='outliers', fig.height=3, fig.cap="The temperature in Melbourne on 1 February 2000, and the following ten days. Ten red lines (two groups of five lines each) mark the times of the outliers. The first group spans 14:30--17:00 on 3 February, while the second group spans 15:30--18:00 on the following day."}
elec %>%
  filter(Date >= "2000-02-01", Date <= "2000-02-11") %>%
  ggplot(aes(x=DateTime, y=Temp)) +
  geom_line() +
  geom_vline(data=outliers, aes(xintercept = DateTime), col='red') +
  labs(y="Temperature", x="Date") +
  ggplot2::theme_bw()
```

>>>>

---
title: "STR: Seasonal-Trend decomposition using Regression"
author:
 - familyname: Dokumentov
   othernames: Alexander
   address: Growing Data\newline Australia
 - familyname: Hyndman
   othernames: Rob J
   address: Monash University\newline Australia
   email: Rob.Hyndman@monash.edu
   correspondingauthor: true
abstract: "We propose a new method for decomposing seasonal data: STR (a Seasonal-Trend decomposition using Regression). Unlike other decomposition methods, STR allows for multiple seasonal and cyclic components, covariates, seasonal patterns that may have non-integer periods, and seasonality with complex topology. It can be used for time series with any regular time index including hourly, daily, weekly, monthly or quarterly data. It is competitive with existing methods when they exist, but tackles many more decomposition problem than other methods allow.\\par STR is based on a regularized optimization, and so is somewhat related to ridge regression. Because it is based on a statistical model, we can easily compute confidence intervals for components, something that is not possible with most existing decomposition methods (such as STL, X-12-ARIMA, SEATS-TRAMO, etc.).\\par
  Our model is implemented in the R package *stR*, so can be applied by anyone  to their own data."
keywords: "time series decomposition, seasonal data, Tikhonov regularisation, ridge regression, LASSO, STL, TBATS, X-12-ARIMA, BSM"
blind: false
cover: false
toc: false
bibliography: strrefs.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    toc: no
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
header-includes:
  - \graphicspath{{figs/}}
  - \usepackage{amsmath}
  - \setlength{\arraycolsep}{3pt}
  - \setcounter{MaxMatrixCols}{20}
  - \newcommand{\argmin}{\operatorname{arg\,min}}
  - \newcommand{\sn}{\kappa}
  - \renewcommand{\vec}{\operatorname{vec}}
  - \newcommand{\diag}{\operatorname{diag}}
  - \renewcommand{\mod}{\operatorname{mod}}
  - \usepackage[section]{placeins}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE
)
library(tidyverse)
library(stR)
library(fabletools)
library(distributional)
library(lubridate)
library(knitr)
library(kableExtra)
set.seed(2020 - 03 - 28)

# Create predictor dist object
predictor_dist <- function(x, level) {
  dist_percentile(
    with(x, mapply(c, data, split(lower, row(lower)), split(upper, row(upper)), SIMPLIFY = FALSE)),
    list(c(50, 50 + rep(c(-1, 1), each = length(level)) * rep(level, 2) / 2))
  )
}

source("simulations.R")
```

@@

# Simulations

To demonstrate that our method estimates the appropriate time series components, we use simulations where we know the true underlying components. We will compare our results against two other decomposition methods that handle multiple seasonal components, STL and TBATS. STL [@cleveland1990stl] is a well known decomposition method, and therefore a natural choice to use as a method to compare. The original methodology and implementation was designed for only one seasonal period. However, the implementation in the forecast package for R [@Rforecast] will compute an STL decomposition iteratively with multiple seasonal periods. TBATS [@de2011forecasting] is a forecasting method designed for multiple seasonalities; it provides a decomposition of the time series into components, and so can provide an alternative comparison.

Our simulated daily data consists of four components:
\begin{equation}
  \label{STR:many_parts2}
  y_{t} = T_{t} + \alpha S^{W}_{t} + \beta S^{Y}_{t} + \gamma R_{t} , \quad t=1,\dots,n,
\end{equation}
where $T_{t}$ is a trend, $S^{W}_{t}$ is a weekly seasonal component, $S^{Y}_{t}$ is a yearly seasonal component, $R_{t}$ is the remainder, and $\alpha$, $\beta$, and $\gamma$ are parameters which control the contribution of the components to $y_{t}$.

We use two data generating processes (DGPs) to simulate data. For the first DGP, we use a quadratic trend function with random coefficients, such that $T_t^* = N_1(t + n/2(N_2-1))^2$ where $N_1$ and $N_2$ are independent N(0,1) random variables. Then $T_t^*$ is normalized to give $T_t$ with mean zero and unit variance. The seasonal components consist of five pairs of Fourier terms with random N(0,1) coefficients, which are then also normalized. Because this DGP has deterministic components, we refer to it as the "Deterministic" GDP.

For the second "Stochastic" DGP, the trend is given by an ARIMA(0,2,0) model with standard normal errors, which is then normalized. The seasonal components are also smooth and obtained using a similar process constrained to be periodic. Specifically, for each seasonal component we start with a vector of iid N(0,1) values of the length of one season, normalized and replicated to be of length $n$. This is then integrated and normalized twice to make it smooth. So each seasonal component is as smooth as the trend, is centred on zero and is periodic.

Figure \@ref(fig:examples) shows examples of simulated time series from each of the two DGPs with $\alpha=\beta=1$ and $\gamma=0.25$. We generated 1095 days in each case (three non-leap years of daily data).

```{r examples, fig.cap="Examples of time series generated by the two DGPs.", fig.height=3.5}
bind_rows(
  get_msts_data(type = "stochastic", l = 1095, gamma=.4) %>% 
    mutate(type = "Stochastic", day = row_number()),
  get_msts_data(type = "deterministic", l = 1095, gamma=.4) %>% 
     mutate(type = "Deterministic", day = row_number())
) %>%
  ggplot(aes(x = day, y = data, group = type)) +
  geom_line() +
  facet_grid(type ~ .)
```

For comparison of STR, STL and TBATS we use parameter sets $(\alpha, \beta, \gamma) = \{(1, 1, 0.2), (1, 1, 0.4), (1, 1, 0.6)\}$, and create 20 time series for each DGP and parameter set. We apply STR, STL and TBATS to each series and compute MSE for each component (averaging across all data sets). The results are shown in Table \@ref(tab:cbetable). It is clear that STR outperforms STL and TBATS in these scenarios.

```{r errors}
n <- 100
errors <- bind_rows(
  compute_n_errors(n = n, gamma = 0.2, type = "stochastic"),
  compute_n_errors(n = n, gamma = 0.4, type = "stochastic"),
  compute_n_errors(n = n, gamma = 0.6, type = "stochastic"),
  compute_n_errors(n = n, gamma = 0.2, type = "deterministic"),
  compute_n_errors(n = n, gamma = 0.4, type = "deterministic"),
  compute_n_errors(n = n, gamma = 0.6, type = "deterministic")
)
```

```{r signif}
errors <- errors %>%
       mutate(method = factor(method, c("STR","STL","TBATS")))

pvalues <- bind_rows(
    get_pvalues(errors, type="stochastic", gamma=0.2),
    get_pvalues(errors, type="stochastic", gamma=0.4),
    get_pvalues(errors, type="stochastic", gamma=0.6),
    get_pvalues(errors, type="deterministic", gamma=0.2),
    get_pvalues(errors, type="deterministic", gamma=0.4),
    get_pvalues(errors, type="deterministic", gamma=0.6)
  ) %>%
  mutate(method = str_remove(term,"method")) %>%
  select(-term) 
```

```{r cbetable}
rmse <- errors %>%
  group_by(type, gamma, method) %>%
  summarise(
    trend = sqrt(mean(trend^2)),
    season_week = sqrt(mean(season_week^2)),
    season_year = sqrt(mean(season_year^2)),
    remainder = sqrt(mean(remainder^2)),
  ) %>%
  ungroup()
rmse %>%
  pivot_longer(trend:remainder, names_to="var", values_to="rmse") %>%
  left_join(pvalues, by=c("type","gamma","method","var")) %>%
  mutate(
    method = factor(method, levels = c("STL","TBATS","STR")),
    rmse = sprintf("%.4f", rmse),
    rmse = if_else(p.value < 0.05, paste0("\\textbf{",rmse,"}"), rmse, missing=rmse)
  ) %>%
  pivot_wider(c(type,gamma,method),names_from="var",values_from="rmse") %>%
  arrange(type,gamma,method) %>%
  select(gamma, method, everything(), -type) %>%
  knitr::kable(
    col.names = c("$\\gamma$", "Method", "Trend RMSE", "Weekly RMSE", "Yearly RMSE", "Remainder RMSE"),
    caption = "RMSE over 100 simulations for each DGP and value of $\\gamma$. Bold values indicate results that are significantly different from the STR values.",
    digits = 3,
    escape = FALSE,
    align = "rlrrrr",
    booktabs = TRUE,
    format = "latex",
    table.envir = "table",
    position = "!ht"
  ) %>%
  kable_styling() %>%
  pack_rows("Stochastic DGP", 1, 9) %>%
  pack_rows("Deterministic GDP", 10, 18) %>%
  row_spec(seq(3, 15, by = 3), extra_latex_after = "\\midrule")
```

@@

```{r supermarket, fig.height=7, fig.cap="Supermarket and grocery store log revenue in New South Wales, decomposed with STR. The original data (logged) are shown in the top panel, while the components are shown in the other panels, with 95\\% confidence intervals for the trend and season components in blue"}
supermarket_nsw <- tsibbledata::aus_retail %>%
  filter(
    State == "New South Wales",
    Industry == "Supermarket and grocery stores",
    year(Month) >= 2000,
    year(Month) <= 2009
  ) %>%
  pull(Turnover) %>%
  ts(start = 2000, frequency = 12)

decomp <- log(supermarket_nsw) %>%
  AutoSTR(confidence = 0.95, gapCV = 12, reltol = 0.00001)

as_tsibble(decomp$input$data) %>%
  rename(log_Data = value) %>%
  mutate(
    Remainder = decomp$output$random$data,
    Trend = predictor_dist(decomp$output$predictors[[1]], 95),
    Season = predictor_dist(decomp$output$predictors[[2]], 95)
  ) %>%
  as_dable(response = "log_Data", aliases = rlang::exprs(log_Data = Trend + Season + Remainder), method = "STR") %>%
  autoplot(size = 0.5, fill = "blue", level = 95) +
  theme(legend.position = "none") +
  xlab("Month") +
  theme_bw() +
  theme(legend.position = "none")
```

@@

This data set has two seasonal patterns. The first pattern is a weekly seasonal pattern ($m_1=48\times7=336$) that represents the specific demand features that are attributable to a particular day of the week. The second pattern is a daily seasonal pattern ($m_2=48$) with the topology in Figure \ref{fig:topology_pdf}, which allows the model to distinguish between working days and holidays/weekends, and to make transitions between them. The pattern reflects the tendency to have a higher electricity demand during standard working hours and a lower demand at night. It also reflects the tendency to have different demand patterns on working days and holidays/weekends. A longer series would also have an annual seasonal pattern, but with only 115 days, this cannot be distinguished from the trend component.

```{r electricity}
# Read and join all data sets
elec <- read_csv("data/VIC.csv") %>%
  mutate(Date = as.Date(Date, origin = "1899-12-30")) %>%
  left_join(
    read_csv("data/Melbourne.csv") %>%
      mutate(Date = as.Date(Date, origin = "1899-12-30")),
    by = c("Date", "Period")
  ) %>%
  left_join(
    read_csv("data/VIC public holidays.csv", quote = "\"", col_names = "Date") %>%
      mutate(
        Date = lubridate::dmy(Date),
        Holidays = TRUE
      ),
    by = c("Date")
  ) %>%
  replace_na(list(Holidays = FALSE)) %>%
  rename(Demand = OperationalLessIndustrial) %>%
  select(-Industrial) %>%
  mutate(Time = seq_along(Date)) %>%
  filter(Date >= "2000-01-11", Date <= "2000-05-04") %>%
  # Add in day, week and working day info
  mutate(
    Weekday = factor(weekdays(Date, abbreviate = TRUE),
      levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
    ),
    DailySeasonality = Period - 1,
    WeeklySeasonality = (as.integer(Weekday) - 1) * 48 + Period - 1,
    WorkingDay = !Holidays & Weekday %in% c("Mon", "Tue", "Wed", "Thu", "Fri"),
    WDSeasonality = (DailySeasonality + (!WorkingDay) * 100),
    DateTime = as.POSIXct(paste(Date, "00:00:00")) + (Period - 1) * 30 * 60
  )

# Have we already done this?
if (file.exists("data/electricity_str.rds")) {
  elec_str <- read_rds("data/electricity_str.rds")
} else {
  # Set up predictors for STR
  Predictors <- list()
  Predictors$Trend <- list(
    name = "Trend",
    data = rep(1, NROW(elec)),
    times = elec$Time,
    seasons = rep(1, NROW(elec)),
    timeKnots = seq(from = first(elec$Time), to = last(elec$Time), length.out = 116),
    seasonalStructure = list(segments = list(c(0, 1)), sKnots = list(c(1, 0))),
    lambdas = c(1500, 0, 0)
  )
  Predictors$WSeason <- list(
    name = "Weekly seas",
    data = rep(1, NROW(elec)),
    times = elec$Time,
    seasons = elec$WeeklySeasonality,
    timeKnots = seq(from = first(elec$Time), to = last(elec$Time), length.out = 12),
    seasonalStructure = list(segments = list(c(0, 336)), sKnots = c(as.list(seq(4, 332, 4)), list(c(336, 0)))),
    lambdas = c(0.8, 0.6, 100)
  )
  Predictors$WDSeason <- list(
    name = "Daily seas",
    data = rep(1, NROW(elec)),
    times = elec$Time,
    seasons = elec$WDSeasonality,
    timeKnots = seq(from = first(elec$Time), to = last(elec$Time), length.out = 24),
    seasonalStructure = list(
      segments = list(c(0, 48), c(100, 148)),
      sKnots = c(as.list(c(1:47, 101:147)), list(c(0, 48, 100, 148)))
    ),
    lambdas = c(0.003, 0, 240)
  )
  Predictors$TrendTempM <- list(
    name = "Trend temp",
    data = elec$Temp,
    times = elec$Time,
    seasons = rep(1, NROW(elec)),
    timeKnots = Predictors$Trend$timeKnots,
    seasonalStructure = Predictors$Trend$seasonalStructure,
    lambdas = c(1e7, 0, 0)
  )
  Predictors$TrendTempM2 <- list(
    name = "Trend temp^2",
    data = elec$Temp^2,
    times = elec$Time,
    seasons = rep(1, NROW(elec)),
    timeKnots = Predictors$Trend$timeKnots,
    seasonalStructure = Predictors$Trend$seasonalStructure,
    lambdas = c(1e7, 0, 0)
  )

@@

```{r vic, fig.height=10, fig.cap="Electricity demand in Victoria, decomposed using STR. The original data are plotted in black in the first panel, with the decomposed trend plotted in red. The subsequent panels contain the weekly seasonal pattern; the daily seasonal pattern with complex topology; the effect of temperature in Melbourne, which has a daily seasonal pattern with complex topology; the effect of squared temperatures in Melbourne, which is allowed to vary over time; and the residuals. Grey vertical lines represent the residuals with the largest absolute values.", dependson='outliers'}
elec_dbl <- elec %>%
  as_tsibble(index = DateTime) %>%
  mutate(
    Remainder = elec_str$output$random$data,
    Trend = predictor_dist(elec_str$output$predictors[[1]], 95),
    Weekly = predictor_dist(elec_str$output$predictors[[2]], 95),
    Daily = predictor_dist(elec_str$output$predictors[[3]], 95),
    Temp = predictor_dist(elec_str$output$predictors[[4]], 95),
    Tempsq = predictor_dist(elec_str$output$predictors[[5]], 95),
  ) %>%
  select(DateTime, Demand, Trend, Weekly, Daily, Temp, Tempsq, Remainder) %>%
  as_dable(
    response = "Demand",
    aliases = rlang::exprs(Demand = Trend + Weekly + Daily + Temp + Tempsq + Remainder),
    method = "STR"
  )

p <- elec_dbl %>%
  autoplot(size = 0.5, fill = "blue", level = 95) +
  xlab("Month") +
  geom_vline(data = outliers, aes(xintercept = DateTime), col = "gray") +
  ggplot2::theme_bw() + theme(legend.position = "none")
# Put vline behind data
p$layers <- p$layers[c(1, 4, 2, 3)]
p
```

@@

```{r outlierstable, results='asis'}
outliers %>%
  mutate(
    Hour1 = hour(DateTime),
    Hour2 = hour(DateTime + 1800),
    Minute1 = minute(DateTime) %>% as.character(),
    Minute2 = minute(DateTime + 1800) %>% as.character(),
    Minute1 = if_else(Minute1 == "0", "00", Minute1),
    Minute2 = if_else(Minute2 == "0", "00", Minute2),
    Time = paste0(Hour1, ":", Minute1, " -- ", Hour2, ":", Minute2),
    Weekday = case_when(
      Weekday == "Thu" ~ "Thursday",
      Weekday == "Fri" ~ "Friday"
    ),
    Date = paste(day(Date), month(Date, label = TRUE, abbr = FALSE), year(Date)),
  ) %>%
  select(Date, Weekday, Time, res) %>%
  knitr::kable(
    format='latex',
    col.names = c("Date", "Day of week", "Time period", "Residual"),
    digits = 1, align = "rlcr", booktabs = TRUE, linesep=10,
    caption = "The ten residuals that are largest in absolute value after an STR decomposition."
  )
```

```{r MelTemp, dependson='outliers', fig.height=3, fig.cap="The temperature in Melbourne on 1 February 2000, and the following ten days. Ten red lines (two groups of five lines each) mark the times of the outliers. The first group spans 14:30--17:00 on 3 February, while the second group spans 15:30--18:00 on the following day."}
elec %>%
  filter(Date >= "2000-02-01", Date <= "2000-02-11") %>%
  ggplot(aes(x = DateTime, y = Temp)) +
  geom_line() +
  geom_vline(data = outliers, aes(xintercept = DateTime), col = "red") +
  labs(y = "Temperature", x = "Date") +
  ggplot2::theme_bw()
```
